### 2019-09-01 paper/paper.tex
Editing.

\section*{ABSTRACT}
%
Previously, we introduced a method of optimizing four primary geometric
parameters of a bicycle's design to minimize its Handling Quality Metric. Here
we expand that method to optimize 23 geometric and inertial parameters in the
benchmark parameterimization of the linear Whipple-Carvallo bicycle model. To
ensure physically realizable bicycle designs we include 7 equality constraints,
21 inequality constraints, and maximal and minamal bounds on each free
opmization parameter. This improves over the prior work by expanding the search
space with many more parameters and the guarntee of realizability. We close by
presenting five bicycle designs that have optimal later handling qualities. The
bicycles are not generally self-stable and have some ununusal characteristics.
%
\begin{keywords}
  bicycle,
  dynamics,
  handling qualities,
  control,
  optimization,
  design
\end{keywords}

\section{INTRODUCTION}
%
Physical design features of ground vehicles can affect their lateral handling
qualities. Geometry, mass, and mass distribution of the vehicle's components as
well as tire characteristics are primary contributors to poor and good handling
due to their important influence on the vehicle's dynamics. In past work, we
have presented a theoretical and computational framework for assessing the
lateral task-independent handling qualities of simplified single track vehicle
designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed that
minimizing our proposed handling quality metric (HQM) can produce theoretically
optimal handling designs when only four geometric parameters are explored as
the optimization variables~\cite{Moore2016}. The present work's goal is to
expand this optimization problem to all of the geometry, mass, and inertial
parameters present in the linear Whipple-Carvallo bicycle
model~\cite{Meijaard2007}.  This broadens the search space considerably but we
constrain it so that only realizable optimal bicycle designs are discovered. To
do so, we formulate a constrained optimization problem and use derivative-free
optimization to discover optimal, yet realizable, bicycle designs.

@@

reflects both a reformulation of and addition to the benchmark parameterization
of the linear Whipple-Carvallo bicycle model~\cite{Meijaard2007}. We call this
the ``principal parameterization'' as opposed to the ``benchmark
parameterization'' in~\cite{Meijaard2007}. This parameterization differs from
the benchmark parameterization in three ways. Firstly, the person and rear
frame are treated as separate rigid bodies each with their on inertial
parameters. Secondly, we express the inertial parameters of each rigid body in
terms of central principal radii of gyration to decouple the mass from the
inertia terms.  Lastly, we introduce two simple dimensional parameters that
define the geometric extents of the person which are used to constrain the
location of the person's body. Table~\ref{tab:parameters} provides the
parameters and the reference values which are derived from the measurements of
a Batavus Browser Bicycle and the rider ``Jason'' presented
in~\cite{Moore2012}. This parameterization can be transformed into the
benchmark parameterization readily, but not vice versa.

@@

    properties of the five rigid bodies: front wheel (orange), rear wheel

@@

    radii of gyration of each rigid body. The primary principal angles,
    $\alpha_{D,P,H}$, are defined as the angle about the $y$ axis from $x$ to
    the maximum principal axes in the $XZ$ plane.}

@@

    and rider ``Jason''. Note that the benchmark parameters are also used in
    the paper and are defined in~\cite{Meijaard2007}.}

@@

    $c$       &  0.068581           & \si{\meter} & Trail \\
    $w$       &  1.1210             & \si{\meter} & Wheelbase \\

@@

    $r_R$     & 0.34096 & \si{\meter}  & Radius \\
    $x_R$     & 0.0 & \si{\meter}      & $X$ mass center coordinate \\
    $y_R$     & 0.0 & \si{\meter}      & $Y$ mass center coordinate \\
    $z_R$     & -0.34096 & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Raa}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $A_R$ \\
    $k_{Rbb}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $B_R$ \\
    $k_{Ryy}$ & 0.22136 & \si{\meter}  & Central principal radii of gyration about $Y$ \\

@@

    $r_F$     & 0.34353 & \si{\meter} & Radius \\
    $x_F$     & 1.1210 & \si{\meter} & $X$ mass center coordinate \\
    $y_F$     & 0.0 & \si{\meter} & $Y$ mass center coordinate \\
    $z_F$     & -0.34353 & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $A_F$ \\
    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $B_F$ \\
    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radii of gyration about $Y$ \\

@@

    $m_P$      &  83.5    & \si{\kilogram} & Mass \\
    $x_P$      &  0.31577 & \si{\meter} & $X$ mass center coordinate \\
    $y_P$      & 0.0      & \si{\meter} & $Y$ mass center coordinate \\
    $z_P$      & -1.0990  & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Paa}$  &  0.36797 & \si{\meter} & Central principal radii of gyration about $A_P$ \\
    $k_{Pbb}$  &  0.15276 & \si{\meter} & Central principal radii of gyration about $B_P$ \\
    $k_{Pyy}$  &  0.36717 & \si{\meter} & Central principal radii of gyration about $Y$ \\
    $\alpha_P$ & 0.18618 & \si{\radian} & Angle about $Y$ between $A_P$ and $X$ \\

@@

    $m_H$ & 3.2200 & \si{\kilogram}     & Mass \\
    $x_H$ & 0.86695 & \si{\meter}       & $X$ mass center coordinate \\
    $y_H$ & 0.0 & \si{\meter}           & $Y$ mass center coordinate \\
    $z_H$ & -0.74824 & \si{\meter}      & $Z$ mass center coordinate \\
    $k_{Haa}$ & 0.29556 & \si{\meter}   & Central principal radii of gyration about $A_H$ \\
    $k_{Hbb}$ & 0.14493 & \si{\meter}   & Central principal radii of gyration about $B_H$ \\
    $k_{Hyy}$ & 0.27630 & \si{\meter}   & Central principal radii of gyration about $Y$ \\
    $\alpha_H$ & 0.36995 & \si{\radian} & Angle about $Y$ between $A_H$ and $X$ \\

@@

    $m_D$ &  9.8600 & \si{\kilogram}    & Mass \\
    $x_D$ &  0.27595 & \si{\meter}      & $X$ mass center coordinate \\
    $y_D$ & 0 & \si{\meter}             & $Y$ mass center coordinate \\
    $z_D$ & -0.53784 & \si{\meter}      & $Z$ mass center coordinate \\
    $k_{Daa}$ &  0.28587 & \si{\meter}  & Central principal radii of gyration about $A_D$ \\
    $k_{Dbb}$ &  0.22079 & \si{\meter}  & Central principal radii of gyration about $B_D$ \\
    $k_{Dyy}$ &  0.36539 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
    $\alpha_D$ &  1.1722 & \si{\radian} & Angle about $Y$ between $A_D$ and $X$ \\

@@

\section*{BOUNDS AND CONSTRAINTS}
%
The optimal principal parameters are subject to a set of constraints designed
to ensure that a physically realizable bicycle is obtained from the
optimization procedure. These constraints are made up of bounds on the free
parameters and both equality and inequality constraints among the parameters.
Below the basic constraint concepts presented and grouped by the associated
rigid body or collection thereof:
%
\begin{description}
  \item[Total $T$] The combination of the five rigid bodies.
    \begin{itemize}
      \item The likely physical extents of the rigid bodies must exist above
        the ground plane.
      \item Both bicycle and rider are symmetric about the rider's sagittal
        plane.
      \item The total mass is below a reasonably human lift-able amount.
      \item The wheels cannot overlap.
      \item The bicycle cannot topple forward during hard braking or backward
        during hard acceleration.
      \item The closed-loop path tracking controlled system~\cite{Hess2012}
        must be stable. This is required to obtain a valid HQM value.
    \end{itemize}

  \item[Rider $P$] A single rigid body represents the rider.
    \begin{itemize}
      \item Rider mass is that of an average person.
      \item The rider's joint angles are fixed in a nominal configuration
        typical of upright bicycling and the resulting mass distribution is
        derived from standard body segment estimation methods.
      \item The rider cannot penetrate the ground.
    \end{itemize}

@@

    Min & & Parameter & & Max\\

@@

    $0.25   $ & $\leq$ & $m_H     $  &  $\leq$  &  $\infty$ \\

@@

To prevent the rider from being positioned and oriented such that their body is
not penetrating the ground we introduce two dimensions that define a cross
whose apex is at the center of mass of the person and the cross axes are
parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
along the principal axis to the tip of the toes and $w_P / 2$ is the distance
along the second principal axes to the tip of the hands. The constraints
$c_2,\ldots,c_5$ are derived from these rules.

>>>>

\section*{ABSTRACT}
%
Previously, we introduced a method of optimizing four primary geometric
parameters of a bicycle's design to maximize its lateral handling qualities.
Here we expand that method to optimize over all of the geometric and inertial
parameters in the linear Whipple-Carvallo bicycle model. To ensure physically
realizable bicycle designs we include 7 equality constraints, 21 inequality
constraints, and maximal and minimal bounds on each free optimization
parameter. This improves over the prior work by expanding the search space with
many more parameters and the guarantee of physical realizability. We present
four bicycle designs discovered by the optimization procedure that have optimal
later handling qualities. The bicycles are similar in design to familiar
bicycle designs but are not generally self-stable and exhibit unusual
characteristics such as large positive and negative trail, large size relative
to the rider, and minimal steering inertia. The method is a useful tool for
generating atypical bicycle designs that exhibit desired dynamical qualities
and could be broadly applied to other vehicle designs.
%
\begin{keywords}
  bicycle,
  dynamics,
  handling qualities,
  control,
  optimization,
  design
\end{keywords}

\section{INTRODUCTION}
%
Physical design features of ground vehicles can affect their lateral handling
qualities. Geometry, mass, and mass distribution of the vehicle's components as
well as tire characteristics are primary contributors to poor and good handling
due to their important influence on the vehicle's dynamics. In past work, we
have presented a theoretical and computational framework for assessing the
lateral task-independent handling qualities of simplified single track vehicle
designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed that
minimizing our proposed handling quality metric (HQM) can produce theoretically
optimal handling designs when only four geometric parameters are explored as
the optimization variables~\cite{Moore2016}. The present work's goal is to
expand this optimization problem to all of the geometry, mass, and inertial
parameters present in the linear Whipple-Carvallo bicycle
model~\cite{Meijaard2007}. This broadens the search space considerably but we
constrain it so that only realizable optimal bicycle designs are discovered. To
do so, we formulate a constrained optimization problem and use derivative-free
optimization to discover optimal, yet realizable, bicycle designs. We close by
discussing the characteristics of the discovered designs.

@@

reflects both a reformulation of and extension to the benchmark
parameterization of the linear Whipple-Carvallo bicycle
model~\cite{Meijaard2007}. We call this the ``principal parameterization'' as
opposed to the ``benchmark parameterization'' in~\cite{Meijaard2007}. This
parameterization differs from the benchmark parameterization in three ways.
Firstly, the person and rear frame are treated as separate rigid bodies each
with their own inertial parameters. Secondly, we express the inertial
parameters of each rigid body in terms of central principal radii of gyration
to decouple the mass from the inertia terms. Lastly, we introduce two simple
dimensional parameters that define the geometric extents of the person which
are used to constrain the location of the person's body in the optimization
problem. Table~\ref{tab:parameters} provides the parameters and the reference
values which are derived from the measurements of a Batavus Browser Bicycle and
the rider ``Jason'' presented in~\cite{Moore2012}. This represents a typical
bicycle and rider. The principal parameterization can be transformed into the
benchmark parameterization readily, but not vice versa. Note that the benchmark
parameter variables names are also used in the paper for convenience and are
defined in~\cite{Meijaard2007}.

@@

    properties of the five rigid bodies, front wheel (orange), rear wheel

@@

    radii of gyration of each rigid body as colored solid lines. The primary
    principal angles, $\alpha_{D,P,H}$, are defined as the angle about the $y$
    axis from $x$ to the maximum principal axes in the $XZ$ plane.}

@@

    and rider ``Jason''. The $A$ and $B$ axes are the maximal and minimal
    principal axes in the $XZ$ plane, respectively.}

@@

    $c$       &  0.069           & \si{\meter} & Trail \\
    $w$       &  1.121             & \si{\meter} & Wheelbase \\

@@

    $r_R$     & 0.341 & \si{\meter}  & Radius \\
    $x_R$     & 0.000 & \si{\meter}      & $X$ mass center coordinate \\
    $y_R$     & 0.000 & \si{\meter}      & $Y$ mass center coordinate \\
    $z_R$     & -0.341 & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Raa}$ & 0.171 & \si{\meter}  & Central principal radii of gyration about $A_R$ \\
    $k_{Rbb}$ & 0.171 & \si{\meter}  & Central principal radii of gyration about $B_R$ \\
    $k_{Ryy}$ & 0.221 & \si{\meter}  & Central principal radii of gyration about $Y$ \\

@@

    $r_F$     & 0.344 & \si{\meter} & Radius \\
    $x_F$     & 1.121 & \si{\meter} & $X$ mass center coordinate \\
    $y_F$     & 0.000 & \si{\meter} & $Y$ mass center coordinate \\
    $z_F$     & -0.344 & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Faa}$ &  0.209 & \si{\meter} & Central principal radii of gyration about $A_F$ \\
    $k_{Fbb}$ &  0.209 & \si{\meter} & Central principal radii of gyration about $B_F$ \\
    $k_{Fyy}$ &  0.272 & \si{\meter} & Central principal radii of gyration about $Y$ \\

@@

    $m_P$      &  83.50    & \si{\kilogram} & Mass \\
    $x_P$      &  0.316 & \si{\meter} & $X$ mass center coordinate \\
    $y_P$      & 0.000      & \si{\meter} & $Y$ mass center coordinate \\
    $z_P$      & -1.099  & \si{\meter} & $Z$ mass center coordinate \\
    $k_{Paa}$  &  0.368 & \si{\meter} & Central principal radii of gyration about $A_P$ \\
    $k_{Pbb}$  &  0.153 & \si{\meter} & Central principal radii of gyration about $B_P$ \\
    $k_{Pyy}$  &  0.367 & \si{\meter} & Central principal radii of gyration about $Y$ \\
    $\alpha_P$ & 0.18618 & \si{\radian} & Angle about $Y$ between $X$ and $A_P$ \\

@@

    $m_H$ & 3.22 & \si{\kilogram}     & Mass \\
    $x_H$ & 0.867 & \si{\meter}       & $X$ mass center coordinate \\
    $y_H$ & 0.000 & \si{\meter}           & $Y$ mass center coordinate \\
    $z_H$ & -0.748 & \si{\meter}      & $Z$ mass center coordinate \\
    $k_{Haa}$ & 0.296 & \si{\meter}   & Central principal radii of gyration about $A_H$ \\
    $k_{Hbb}$ & 0.145 & \si{\meter}   & Central principal radii of gyration about $B_H$ \\
    $k_{Hyy}$ & 0.276 & \si{\meter}   & Central principal radii of gyration about $Y$ \\
    $\alpha_H$ & 0.36995 & \si{\radian} & Angle about $Y$ between $X$ and $A_H$ \\

@@

    $m_D$ &  9.86 & \si{\kilogram}    & Mass \\
    $x_D$ &  0.276 & \si{\meter}      & $X$ mass center coordinate \\
    $y_D$ & 0.000 & \si{\meter}             & $Y$ mass center coordinate \\
    $z_D$ & -0.538 & \si{\meter}      & $Z$ mass center coordinate \\
    $k_{Daa}$ &  0.286 & \si{\meter}  & Central principal radii of gyration about $A_D$ \\
    $k_{Dbb}$ &  0.221 & \si{\meter}  & Central principal radii of gyration about $B_D$ \\
    $k_{Dyy}$ &  0.365 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
    $\alpha_D$ &  1.1722 & \si{\radian} & Angle about $Y$ between $X$ and $A_D$ \\

@@

\section{BOUNDS AND CONSTRAINTS}
%
The optimal principal parameters are subject to a set of constraints designed
to ensure that a physically realizable bicycle is obtained from the
optimization procedure. These constraints are made up of bounds on the free
parameters and both equality and inequality constraints among the parameters.
Below the basic constraint concepts presented and grouped by the associated
rigid body or collection thereof:
%
\begin{description}
  \item[Total $T$] The combination of the five rigid bodies.
    \begin{itemize}
      \item The likely physical extents of the rigid bodies must exist above
        the ground plane.
      \item Both bicycle and rider are symmetric about the rider's sagittal
        plane.
      \item The total mass is below a reasonably human lift-able amount.
      \item The wheels cannot overlap.
      \item The bicycle cannot topple forward during hard braking or backward
        during hard acceleration.
      \item The closed-loop path tracking controlled system~\cite{Hess2012}
        must be stable. This is required to obtain a valid HQM value.
    \end{itemize}

  \item[Person $P$] A single rigid body represents the rider.
    \begin{itemize}
      \item Rider mass is that of a typical person.
      \item The rider's joint angles are fixed in a nominal configuration
        typical of upright bicycling and the resulting mass distribution is
        derived from standard body segment estimation methods.
      \item The rider cannot penetrate the ground.
    \end{itemize}

@@

    Lower & & Parameter & & Upper \\

@@

    $0.25\si{\kilogram} $ & $\leq$ & $m_H     $  &  $\leq$  &  $\infty$ \\

@@

To prevent the rider from being positioned and oriented such that their body is
penetrating the ground we introduce two dimensions that define a cross whose
apex is at the center of mass of the person and the cross axes are parallel to
the principal axes in the $XZ$ plane. $l_P / 2$ is the distance along the
principal axis to the tip of the toes and $w_P / 2$ is the distance along the
second principal axes to the tip of the hands. The constraints $c_2,\ldots,c_5$
are derived from these rules.
